### 2015-03-25 markdown/manuscript.tex
exposition exposition exposition

\subsection{GP as ``mangle-ish'' practice}\hypertarget{gp-as-mangle-ish-practice}{}\label{gp-as-mangle-ish-practice}

@@

\section{What it can mean to ``use'' genetic programming}\hypertarget{what-it-can-mean-to-use-genetic-programming}{}\label{what-it-can-mean-to-use-genetic-programming}

\section{Cases for and against ``usable GP'': cost, features, reliability, simplicity}\hypertarget{cases-for-and-against-usable-gp-cost-features-reliability-simplicity}{}\label{cases-for-and-against-usable-gp-cost-features-reliability-simplicity}

\section{``Artificial Intelligence investigated by Special Crimes Unit''; or, Even Athena needed the head of Zeus}\hypertarget{artificial-intelligence-investigated-by-special-crimes-unit-or-even-athena-needed-the-head-of-zeus}{}\label{artificial-intelligence-investigated-by-special-crimes-unit-or-even-athena-needed-the-head-of-zeus}

\section{The mangle of practice and the mangle in practice}\hypertarget{the-mangle-of-practice-and-the-mangle-in-practice}{}\label{the-mangle-of-practice-and-the-mangle-in-practice}

\section{The yardstick and the bamboo hand}\hypertarget{the-yardstick-and-the-bamboo-hand}{}\label{the-yardstick-and-the-bamboo-hand}

@@

\section{``Batch'' and ``interactive'' styles}\hypertarget{batch-and-interactive-styles}{}\label{batch-and-interactive-styles}

\section{Expressing ``resistance'': the problem of Bertrand's Paradox}\hypertarget{expressing-resistance-the-problem-of-bertrands-paradox}{}\label{expressing-resistance-the-problem-of-bertrands-paradox}

@@

\section{The deeper question: What should it mean to \emph{act intelligently}?}\hypertarget{the-deeper-question-what-should-it-mean-to-act-intelligently}{}\label{the-deeper-question-what-should-it-mean-to-act-intelligently}

>>>>

\subsection{GP as ``mangle-ish practice''}\hypertarget{gp-as-mangle-ish-practice}{}\label{gp-as-mangle-ish-practice}

@@

[more here]

\section{As if we \emph{really} meant it}\hypertarget{as-if-we-really-meant-it}{}\label{as-if-we-really-meant-it}

In the same way that Braithwaite's onerous coding exercise drives the attention of its participants toward test-driven design with the obligation to write ``real'' code \emph{only as a refactoring}, I will here demand a \emph{warrant} for every step that moves the algorithm away from random guessing. Not only do traditional search operators like crossover, mutation and [negative] selection not come ``for free'' in this variant, but in every case we must develop a cogent, data-driven argument in favor of starting them \emph{as part of an ongoing search process}. Similarly, the initial selection criteria will be limited to a minimal subset of the training data, and expansion (and other alterations) of the training set will have to be made in light of measured progress, not assumptions that ``more is better'' in every case.

The result will be an incremental process of refinement of an ongoing search, carried out not at the level of externally-assigned parameter ``tweaks'' but rather by \emph{opening} the black boxes we typically demand and demanding we do surgery to correct their ``pathologies'' (and understand their mysteries) without killing them outright. It is not intended as an ``algorithm'' to supplant those used today, but rather as a forced re-description of what we actually already do.

\subsection{A tableau representation for GP systems}\hypertarget{a-tableau-representation-for-gp-systems}{}\label{a-tableau-representation-for-gp-systems}

It will help if I present a bit of formalism.

Operators
Individuals
Rubrics

\subsubsection{``traditional GP tableau''}\hypertarget{traditional-gp-tableau}{}\label{traditional-gp-tableau}

The analogy to SQL.

\subsubsection{``lexicase selection tableau''}\hypertarget{lexicase-selection-tableau}{}\label{lexicase-selection-tableau}

The analogy to SQL.

\subsubsection{``lazy lexicase selection tableau''}\hypertarget{lazy-lexicase-selection-tableau}{}\label{lazy-lexicase-selection-tableau}

The analogy to NoSQL.

\subsubsection{``mutually lazy tableau''}\hypertarget{mutually-lazy-tableau}{}\label{mutually-lazy-tableau}

``Send me a message.''

\subsubsection{``as if we really meant it''}\hypertarget{as-if-we-really-meant-it-1}{}\label{as-if-we-really-meant-it-1}

I need your help.

@@

\subsection{dots and lines, in pushforth}\hypertarget{dots-and-lines-in-pushforth}{}\label{dots-and-lines-in-pushforth}

\subsection{the problem of the unknown language}\hypertarget{the-problem-of-the-unknown-language}{}\label{the-problem-of-the-unknown-language}

\subsection{the problem of the objective(s)}\hypertarget{the-problem-of-the-objectives}{}\label{the-problem-of-the-objectives}

\subsection{expanding the goals when it's boring}\hypertarget{expanding-the-goals-when-its-boring}{}\label{expanding-the-goals-when-its-boring}

@@

\section{Final thoughts: What should it mean to \emph{act intelligently}?}\hypertarget{final-thoughts-what-should-it-mean-to-act-intelligently}{}\label{final-thoughts-what-should-it-mean-to-act-intelligently}

(It would mean the sort of self-awareness it takes to notice that something is wrong, and to ask for help.)
